---
title: "R Notebook"
output: html_notebook
---

## PD model Moneyflow

The plan of the analysis:

1.  Environment setup, familiarizing myself with data
2.  Data cleaning and exploration - missing values, vaiable types, distributions
3.  Exploring relationships of variables with the default rate
4.  Setup an XGBoost model, find hyperparameters, check number of variables in the model
5.  Fit final model, explore results

### Setup and first look at data

```{r}
library(readxl) # excel import
library(pillar) # glimpse
library(ggplot2) # visualizations
library(dplyr) # data wrangling
library(lubridate) # dates
library(rlang) # debugging
library(Hmisc) # multi- histogram
library(Information) # Information value
library(xgboost)
library(superml)

invoices <- read_excel("Moneyflow DS interview Case.xlsx", 
                       sheet = "Train")

```

First glance at data

```{r}
dim(invoices)
glimpse(invoices)
```

There are 21 features of which invoice_id won't be used for feature generation. That leaves 20 features to explore, including default. 7 inputs relating to seller, 7 related to payer and 5 to the transaction.

### Cleaning and exploration

```{r}
table(invoices$defaulted, useNA = 'ifany')
table(invoices$seller_is_public, useNA = 'ifany')
table(invoices$seller_is_public, invoices$defaulted, useNA = 'ifany') # remove
table(invoices$seller_danish_pep, useNA = 'ifany')

table(invoices$payer_is_public, useNA = 'ifany')
table(invoices$payer_danish_pep, useNA = 'ifany')
table(invoices$payer_eu_sanctions, useNA = 'ifany') # remove

table(invoices$offer_seller_is_first_invoice, useNA = 'ifany')

nrow(table(invoices$seller_main_industry_code))
nrow(table(invoices$seller_company_type))
nrow(table(invoices$payer_main_industry_code))
nrow(table(invoices$payer_company_type))
table(invoices$seller_company_type)

qplot(invoices$invoice_amount)
qplot(invoices$invoice_amount[invoices$invoice_amount<100000])
qplot(invoices$seller_age)
qplot(invoices$payerr_age)
qplot(invoices$seller_risika_score)

qplot(invoices$inv_iss2due)
qplot(invoices$inv_iss2purch)
```

Target is binomial, no missing values.

seller_is_public only has 12 values that are all missing.

Turn T/F variables into numeric.

Turn payer age, seller age, invoice amount, seller risika score, payer risika score to numeric

Find hypothetical date by subtracting invoice_issue_date from current date

Add relative sizes of seller and payer industry size and company type

```{r}
invoices_clean <- invoices %>% 
  select(- invoice_id,
         - seller_is_public #since there are only 12 observations which are all missing
         - payer_eu_sanctions) %>% 
  mutate(
    response = case_when(defaulted == 'y' ~ 1, T ~ 0),
    seller_danish_pep_num = as.numeric(seller_danish_pep),
    payer_is_public_num = suppressWarnings(as.numeric(payer_is_public)),
    payer_danish_pep_num = as.numeric(payer_danish_pep),
    payer_eu_sanctions_num = as.numeric(payer_eu_sanctions),
    
    payer_risika_score_num = suppressWarnings(as.numeric(payer_risika_score)),
    
    offer_seller_is_first_invoice_num = suppressWarnings(as.numeric(offer_seller_is_first_invoice)),
    payer_age_num = suppressWarnings(case_when(as.numeric(payerr_age) == 2021 ~ 1,
                                               T ~ as.numeric(payerr_age))), # assuming 2021 is an error
    seller_age_num = as.numeric(seller_age),
    invoice_issue_date = today() - days(inv_due2today),
    invoice_issue_yearmonth = paste0(year(invoice_issue_date),
                                     case_when(
                                       nchar(month(invoice_issue_date)) == 1 ~ paste0('0',month(invoice_issue_date)), 
                                       T ~ as.character(month(invoice_issue_date)))),
    seller_payer_company_type_match = case_when(seller_company_type == payer_company_type ~ 1, T ~ 0),
    seller_payer_industry_type_match = case_when(seller_main_industry_code == payer_main_industry_code ~ 1, T ~ 0)) %>%
  
  left_join(seller_industry_size, by = 'seller_main_industry_code') %>% 
  left_join(seller_company_size, by = 'seller_company_type') %>% 
  left_join(payer_industry_size, by = 'payer_main_industry_code') %>% 
  left_join(payer_company_size, by = 'payer_company_type') %>% 
  mutate(payer_comp_type_f = case_when(payer_comp_size >0.50 ~'Large',
                                            payer_comp_size > 0.1 ~ 'Medium',
                                            T ~ 'Small'),
         seller_comp_type_f = case_when(seller_comp_size >0.50 ~'Large',
                                       seller_comp_size > 0.1 ~ 'Medium',
                                           T ~ 'Small'),
         payer_industry_size_f = case_when(payer_ind_size >0.1 ~'Large',
                                           payer_ind_size > 0.01 ~ 'Medium',
                                           T ~ 'Small'),
         seller_industry_size_f = case_when(seller_ind_size >0.07 ~'Large',
                                       seller_ind_size > 0.01 ~ 'Medium',
                                       T ~ 'Small'))

hist.data.frame(invoices_clean %>% select_if(is.numeric), n.unique =  2)

invoices_clean %>% 
  group_by(invoice_issue_yearmonth) %>% 
  summarise(nn = n(),
            br = mean(response))
```

### Relationship between variables and response

I will be using Information Value & Weight of Evidence to evaluate how each variable relates to response.

```{r}
infvalue <- create_infotables(data=invoices_clean, y="response", bins=10, parallel=FALSE)

infvalue$Summary %>% View()

cormatrix <- cor(invoices_clean %>% select_if(is.numeric), use = 'pairwise.complete.obs')

round(cormatrix,2)
```

### Grid search

```{r}

initial_set_variables <- c(
  'payer_age_num',
  'payer_ind_size',
  'seller_ind_size',
  'offer_cost_percent',
  'payer_risika_score_num',
  'seller_age_num',
  'invoice_amount',
  'inv_due2today',
  'inv_iss2due',
  'seller_risika_score',
  'seller_danish_pep_num',
  'seller_comp_size',
  'payer_comp_size',
  
  'response'
)

set.seed(101) # Set Seed so that same sample can be reproduced in future also
# Now Selecting 75% of data as sample from total 'n' rows of the data  
sample <- sample.int(n = nrow(invoices_clean), size = floor(.75*nrow(invoices_clean)), replace = F)
train_dat <- invoices_clean[sample, initial_set_variables]
test_dat  <- invoices_clean[-sample,initial_set_variables ]

mean(train_dat$response)
mean(test_dat$response)

xg <- XGBTrainer$new(early_stopping = 50,
                     objective = 'binary:logistic')
xgb_model_1 <- GridSearchCV$new(trainer = xg,
                                parameters = list(n_estimators = c(1500),
                                                  learning_rate = c(0.01, 0.1),
                                                  max_depth = c(5,2,10)),
                                n_folds = 3,
                                scoring = c('accuracy','auc'))


xgb_model_1$fit(train_dat , "response")
xgb_model_1$evaluation_scores
```

```{r}
initial_model.cv <- xgb.cv(data = xgb.DMatrix(train_dat %>%
                                                select(-response) %>% 
                                                as.matrix(), 
                                              label = train_dat$response),
                         params = list(
                           eta = 0.01,
                           max_depth = 10,
                           objective = "binary:logistic",
                           eval_metric = "auc"),
                         nfold = 3,
                         nrounds = 5000)

ggplot(initial_model.cv$evaluation_log,aes(x=iter)) + 
  geom_point(aes(y=test_auc_mean), color = "green") +
  geom_point(aes(y=train_auc_mean), color = "red")

initial_model <- xgboost(data = xgb.DMatrix(train_dat %>%
                                              dplyr::select(- response,
                                                            - inv_due2today) %>%
                                              select_if(is.numeric) %>% 
                                              as.matrix(), 
                                            label = train_dat$response),
                         params = list(
                                    eta = 0.01,
                                    max_depth = 10,
                                    objective = "binary:logistic",
                                    eval_metric = "auc"),
                         
                         nrounds = 1750)

xgb.importance(model=initial_model)  
```

```{r}
adjusted_set_variables <- c(
  'payer_age_num',
  'payer_ind_size',
  'seller_ind_size',
  'offer_cost_percent',
  'payer_risika_score_num',
  'seller_age_num',
  'invoice_amount',
  'inv_due2today',
  'inv_iss2due',
  # 'seller_danish_pep_num',
  # 'seller_comp_size',
  # 'payer_comp_size',
  'seller_risika_score',
  
  'response'
)

tuned_model.cv <- xgb.cv(data = xgb.DMatrix(train_dat %>%
                                                select(all_of(adjusted_set_variables),
                                                       -response) %>% 
                                                as.matrix(), 
                                              label = train_dat$response),
                         params = list(
                           eta = 0.01,
                           max_depth = 10,
                           objective = "binary:logistic",
                           eval_metric = "auc"),
                         nfold = 3,
                         nrounds = 5000)

ggplot(tuned_model.cv$evaluation_log,aes(x=iter)) + 
  geom_point(aes(y=test_auc_mean), color = "green") +
  geom_point(aes(y=train_auc_mean), color = "red")
```

```{r}
tuned_model <- xgboost(data = xgb.DMatrix(train_dat %>%
                                              dplyr::select(all_of(adjusted_set_variables),
                                                            - response) %>%
                                              select_if(is.numeric) %>% 
                                              as.matrix(), 
                                            label = train_dat$response),
                         params = list(
                                    eta = 0.01,
                                    max_depth = 10,
                                    objective = "binary:logistic",
                                    eval_metric = "auc"),
                         
                         nrounds = 730)

xgb.importance(model=tuned_model) 
```

PCA

```{r}
train_pca <- prcomp(train_dat %>% select(adjusted_set_variables), center = T, scale. = T, na.action("na.omit"))
```

Look at test data

```{r}
test_dat$predicted <- predict(tuned_model,
                              xgb.DMatrix(test_dat %>%
                                              dplyr::select(all_of(adjusted_set_variables),
                                                            - response) %>%
                                              select_if(is.numeric) %>% 
                                              as.matrix()),
                              type = 'response')
qplot(test_dat$predicted)

test_dat %>% 
  group_by(prec_groups = cut(predicted, breaks = seq(0,0.05,0.01))) %>% 
  summarise(nn = n(),
            br = mean(response),
            distr = nn/nrow(.))

test_dat %>% 
  group_by(prec_groups = cut(predicted, breaks = seq(0,0.01,0.001))) %>% 
  summarise(nn = n(),
            br = mean(response),
            distr = nn/nrow(.))

Metrics::auc(test_dat$response, test_dat$predicted)
```

Further areas of exploration:

-   Check what happened in that one month

-   Fix is_first_invoice_seller variable

-   check what is wrong with highest payer risika scores

-   
